---
title: "Final Project Analysis"
author: "Matt McCoy, HwiJoon Lee, Tanvi Magdum"
date: "23 April 2024"
output: html_notebook
---

```{r InstallPackages, results='hide', echo=FALSE}
if(!require(dplyr)) install.packages("dplyr")
if(!require(gginference)) install.packages("gginference")
library("ggpubr")
```

### 

```{r, results='hide', echo=FALSE}
survey <- read.csv("dataset/CleanedSurvey.csv", stringsAsFactors = TRUE)
```

### Does the use of code generation impact the academic performance in OOD?

**Null Hypothesis:** The use of code generation does not affect the academic performance in OOD.

**Alternate Hypothesis:** The use of code generation affect the academic performance in OOD.

**Statistical Test: T-Test / Trying to compare means between two groups.**

```{r}
survey$CodeGeneration <- ifelse(survey$How.do.you.use.AI.to.assist.your.work.if.you.use.them. == "Code Generation", "Yes", "No")

survey$What.was.or.is.your.grade.from.OOD. <- as.numeric(as.character(survey$What.was.or.is.your.grade.from.OOD.))

grades_codeGen <- survey[survey$CodeGeneration == "Yes", "What.was.or.is.your.grade.from.OOD."]
grades_noCodeGen <- survey[survey$CodeGeneration == "No", "What.was.or.is.your.grade.from.OOD."]

my_data <- survey[c("CodeGeneration", "What.was.or.is.your.grade.from.OOD.")]
names(my_data) <- c("group", "grades")

ggboxplot(my_data, x = "group", y = "grades",
          color = "group", palette = c("#00AFBB", "#E7B800"),
          order = c("Yes", "No"),
          ylab = "Grades from OOD", xlab = "Use of AI for Code Generation")

shapiro.test(grades_codeGen)
shapiro.test(grades_noCodeGen)

```

```{r}
t.test(grades_codeGen, grades_noCodeGen, var.equal = FALSE)

```

### Does AI effect how much time students spent on Object Oriented Design per week?

**Null Hypothesis:** AI does not effect how much time students spend on Object Oriented Design per week.

**Alternate Hypothesis:** AI does effect how much time students spend on Object Oriented Design per week.

**Statistical Test: T-Test**

***Reasoning:*** Conducting a T-Test for this question will produce the best result as comparing the mean number of hours of people who utilize AI in OOD versus those who do not. Will answer whether or not AI had an effect on overall time spent per week doing assignments for the course.

```{r}
noAiOnHoursSpent <- survey %>% filter(grepl("Did not use", What.kind.of.AI.Model.did.you.use.))
AiOnHoursSpent <- survey %>% filter(!grepl("Did not use", What.kind.of.AI.Model.did.you.use.))

t.test(noAiOnHoursSpent$How.many.hours.did.you.spend.on.OOD.per.week., AiOnHoursSpent$How.many.hours.did.you.spend.on.OOD.per.week.)

ggttest(t.test(noAiOnHoursSpent$How.many.hours.did.you.spend.on.OOD.per.week., AiOnHoursSpent$How.many.hours.did.you.spend.on.OOD.per.week.))
```

***Analysis:***

Since the p-value (0.5007) is greater than 0.05, we fail to reject the null hypothesis. This means that there isnt enough evidence to say that there is a significant effect on time spent per week in OOD if someone is utilizing AI verses someone who is not.

Since the 95 percent confidence interval includes 0 it further supports this conclusion that there isnt a significant effect on time spent if using AI.

In conclusion based on the data and the results of the t-test, we do not have sufficient evidence to conclude that there is a significant effect in time spent per week in Object Oriented Design if you are using AI.

**Determining Outliers**

```{r}
boxplot(noAiOnHoursSpent$How.many.hours.did.you.spend.on.OOD.per.week.)
boxplot(AiOnHoursSpent$How.many.hours.did.you.spend.on.OOD.per.week.)
```

As seen in the above plots there are no visible outliers. So we do not need to do anything further.
